---
title: "359HW2"
author: "Mikayla Mulgrew V00923052"
date: "15/10/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Q1

(a)
```{r Q1}

#a function to demonstrate the CLT wrt the uniform distribution
unifclt<-function(ns,size){
  
  #initialize a vector to hold mean values
  means<-rep(1,size)
  
  #for the given number and size of samples
  for (i in 1:ns){
    
    for (j in 1:size){
      #use runif to generate random data from a uniform distribution
      unDist<-runif(size, min=0, max=1)
      mu<-mean(unDist)
      means[i]<-mu
    }
  }
  #plot means of each generated sample 
  hist(means,breaks=seq(0,1,by=0.05), main=paste('Uniform Distn:',ns,'Samples of Size',size, sep=" "), xlab='Mean', ylab='Frequency')
}

par(mfrow=c(2,2))
unifclt(10,10)
unifclt(10,20)
unifclt(10,50)
unifclt(10,100)
unifclt(100,10)
unifclt(100,20)
unifclt(100,50)
unifclt(100,100)
unifclt(1000,10)
unifclt(1000,20)
unifclt(1000,50)
unifclt(1000,100)


#a function to demonstrate the CLT wrt the poisson distribution
poissonclt<-function(ns, size){
  
  #initialize a vector to hold mean values
  means<-rep(1,size)
  
  for (i in 1:ns){
    
    #use rpois to generate random data from a poisson distribution
    for (j in 1:size){
      pDist<-rpois(size,lambda=5)
      mp<-mean(pDist)
      means[i]<-mp
    }
  }
  #plot means of each generated sample 
  hist(means,breaks=seq(0,10,by=0.5), main=paste('Poisson Distn:',ns,'Samples of Size',size, sep=" "), xlab='Mean', ylab='Frequency')
}

par(mfrow=c(2,2))
poissonclt(10,10)
poissonclt(10,20)
poissonclt(10,50)
poissonclt(10,100)
poissonclt(100,10)
poissonclt(100,20)
poissonclt(100,50)
poissonclt(100,100)
poissonclt(1000,10)
poissonclt(1000,20)
poissonclt(1000,50)
poissonclt(1000,100)



#a function to demonstrate the CLT wrt the binomial distribution
bclt<-function(ns,size){
  
  #initialize a vector to hold mean values
  means<-rep(1,size)
  
  for (i in 1:ns){
    
    for (j in 1:size){
      #use rbinom to generate random data from a binomial distribution
      bDist<-rbinom(size, n=1, p=0.20)
      mb<-mean(bDist)
      means[i]<-mb
    }
  }
  #plot means of each generated sample 
  hist(means, main=paste('Binomial Distn:',ns,'Samples of Size',size, sep=''), xlab='Mean', ylab='Frequency')
}

par(mfrow=c(2,2))
bclt(10,10)
bclt(10,20)
bclt(10,50)
bclt(10,100)
bclt(100,10)
bclt(100,20)
bclt(100,50)
bclt(100,100)
bclt(1000,10)
bclt(1000,20)
bclt(1000,50)
bclt(1000,100)

```
(b)
As sample sizes increase, there is a greater spread of the means across the x-axis. 

(c)
As the number of samples increases, the distributions increasingly resemble a normal distribution. 

(d)
The distribution that the samples are generated from has an effect on how approximately normal the distribution of the sample means will ultimately be. If the samples are pulled from a distribution that is quite different from the normal distribution, then the sample sizes will need to be quite large in order to better approximate the normal distribution. 


##Q2

(a)
```{r Q2}

#read salt.txt
old_salt<-read.table(file="C:/Users/Mik/Documents/STAT359/data/salt.txt", sep="", header=TRUE)

#fix data
salt<-rep(1:length(old_salt$X13.53)+1)

for (i in 1:length(old_salt$X13.53)){
  num=old_salt$X13.53[i]
  salt[i]<-num
}
salt[9]<-13.53

##graphical representations of data
par(mfrow=c(1,2))
hist(salt, main='Histogram of Saltiness',xlab='Measure of Saltiness')
qqnorm(salt)
qqline(salt)
boxplot(salt, main='Measure of Saltiness', col='palevioletred')

```
(b)
```{r Q2(b)}

##calculate skew
skew<- function(s){
  
  sumsCubed<-rep(1:length(s))
  
  for (i in 1:length(s)){
    ms<-mean(s)
    sumsCubed[i]<-(s[i]-ms)^3
  }
  
  sk<-sum(sumsCubed)/(length(sumsCubed)*sqrt(var(salt))^3)
}

print(skew(salt))


##calculate kurtosis
kurtosis<- function(s){
  
  sumsQuad<-rep(1:length(s))
  
  for (i in 1:length(s)){
    ms<-mean(s)
    sumsQuad[i]<-(s[i]-ms)^4
  }
  
  kurt<-((sum(sumsQuad)/length(sumsQuad))/var(salt)^2)-3
}

print(kurtosis(salt))

```
(c)
```{r Q2(c)}

##bootstrapping the salt data
skewBoot<- function(s, r){
  
  #number of bootstrap replicates
  reps<-r
  
  #populate a matrix with the replicates
  saltyboot<-matrix(data=sample(x=s,size=reps*length(s),replace=TRUE),nrow=length(s),ncol=reps)
  
  #apply the skew function saltyboot matrix
  skewsb<-apply(saltyboot,2,skew)
  
  #obtain the 95% C.I.
  sbint<-quantile(skewsb, probs=c(0.025,0.975))
  print(sbint)
}

skewBoot(salt, 15000)

```
It seems that the population generating the data may have a slightly skewed distribution from the graphical observations but since the confidence interval contains 0, it is possible for the data to resemble a normal distribution. With a skewness=0.1723753, the data seems fairly symmetrical. 

(d)
```{r Q2(d)}

kurtBoot<- function(s,r){
  
  #number of bootstrap replicates
  reps<-r
  
  #populate a matrix with the replicates
  saltyboot<-matrix(data=sample(x=s,size=reps*length(s),replace=TRUE),nrow=length(s),ncol=reps)
  
  #apply the skew function saltyboot matrix
  kurtsb<-apply(saltyboot,2,kurtosis)
  
  #obtain the 95% C.I.
  sbint<-quantile(kurtsb, probs=c(0.025,0.975))
  print(sbint)
  
}

kurtBoot(salt, 15000)

```
Since the bootstrap confidence interval for the kurtosis contains 0, it does not seem that the population distribution generating the data would have a non-zero kurtosis. 

(e)
Based on all of the above observations, the distribution of the original sample seems very similar to a normal distribution but slightly positive skewed with heavier tails than normal and a flatter peak. 

The qqplot doesn't indicate linearity but rather a distribution that has heavier tails. The boxplot shows this as well with extreme values on either end. 

The histogram indicates a slightly right skewed distribution. 

The values for skewness and kurtosis, 0.1723753 and -0.9342198, respectively, reflect a a slightly right-skewed distribution with a flatter peak. 

The bootstrap confidence intervals indicate that the distribution may not be drastically different than a normal distribution. 


##Q3

```{r Q3}

eggs<-read.table(file='C:/Users/Mik/Documents/STAT359/data/fecundity.txt', sep="", header=TRUE)

par(mfrow=c(1,2))
qqnorm(eggs$RS, main='RS')
qqline(eggs$RS)
qqnorm(eggs$NS, main='NS')
qqline(eggs$NS)

boxplot(eggs$RS, eggs$NS, col='darkorange', names=c('RS', 'NS'))
title('Per Diem Fecundity')

summary(eggs)

var.test(eggs$RS, eggs$NS)

t.test(eggs$RS, eggs$NS, alternative='two.sided',mu=0, var.equal=TRUE)

t.test(eggs$RS, eggs$NS, alternative='less',mu=0, var.equal=TRUE)

```
Using var.test to test the difference in variances between NS and RS gives a p-value of 0.4974. This p-value isn't significant enough to claim that the variances differ. 

Assuming that the variances are equal, a two sided t.test and a lower one sided t.test offer up significant p-values. In conclusion, the null hypothesis is rejected in favor of either alternative hypothesis which suggests a difference in population mean fecundity. 

##Q4

(a)
```{r Q4(a)}

fabrics<-read.csv('C:/Users/Mik/Documents/STAT359/data/fabric2.csv',stringsAsFactors=FALSE, fileEncoding="UTF-8-BOM", header=TRUE)

hqf<-fabrics$H[!is.na(fabrics$H)]
pqf<-fabrics$P[!is.na(fabrics$P)]

par(mfrow=c(1,2))
qqnorm(hqf)
qqline(hqf)
qqnorm(pqf)
qqline(pqf)

```
(b)
```{r Q4(b)}

boxplot(hqf, pqf,col='cadetblue3', names=c('High Quality Fabric', 'Poor Quality Fabric'),main='High-Quality and Poor-Quality Fabrics')

```
The boxplots for High-Quality Fabric and Poor-Quality Fabric overlap significantly but the High-Quality Fabric has a longer tail in the lower values and the Poor-Quality Fabric has a longer tail in the higher values. The averages for each seem quite close to each other so this suggests that there is not a large difference between true average extensibility for the two fabrics. 

(c)
```{r Q4(c)}

var.test(hqf, pqf)

t.test(hqf, pqf, alternative='two.sided', mu=0, var.equal=TRUE)

```
There wasn't sufficient evidence suggesting that the population variances were unequal so a t-test assuming their equality was performed. Given a p-value of 0.6801, we do not reject the null hypothesis and assume that it is likely there is no difference in the average extensibility of the fabrics. 